NHF1 Tandem Deamination Data Generation

### 2022-07-30 R/5'CutSiteAnalysisDataGeneration.rmd
HYV is better than YR!
Lots of changes to accommodate the change in sequence pattern from YR to HYV:
- Changed sequence frequency generation to except regular expression patterns.
- Updated various plotting and data generation notebooks to generate new HYV plots. This also included updating figure generation.

deaminationDataDirectory = file.path(bioinformaticsDirectory, "deamination_determination", "data")
NHF1GeneralReadSequenceDataDirectory = file.path(deaminationDataDirectory, "NHF1", "general_read_sequences")
GM12878GeneralReadSequenceDataDirectory = file.path(deaminationDataDirectory, "GM12878", "general_read_sequences")
DmGeneralReadSequenceDataDirectory = file.path(deaminationDataDirectory, "Dm", "general_read_sequences")

humanReadMinMax = c(22, 30)
drosophilaReadMinMax = c(23, 32)
```

@@

```{r define_all_that_and_a_bag_of_chips_function}
cutSiteAnalysis = function(lesion, repetition, timepoint, cellType, directory,
                           querySequences, sequencesName, readMinMax) {
  readData = fread(generateFilePath(
    directory, dataTypeStrings$alignedReads, ".bed",
    cellType = cellType, lesion = lesion, timepoint = timepoint,
    repitition = repetition, expansionNum = expansionOffset, filtering = filtering
  ))
  readData = data.table(Read_Length = nchar(readData$V7)-expansionOffset*2, Read_Sequence = readData$V7)
  readData = readData[Read_Length >= readMinMax[1] & Read_Length <= readMinMax[2]]

  seqFreq = getSequenceFreqByReadLengthAndPos(readData, querySequences = querySequences)
  fwrite(seqFreq, sep = '\t', file = generateFilePath(
      directory = directory, dataTypeStrings$sequenceFrequency, ".tsv",
      cellType = cellType, lesion = lesion, repitition = repetition,
      timepoint = timepoint, sequence = sequencesName, filtering = filtering))
}
```

```{r compute_and_write_sequence_frequencies}
cutSiteAnalysis("CPD", "all_reps", "1h", "NHF1", NHF1GeneralReadSequenceDataDirectory,
                c("CG","CA","TG","TA"), "YR", humanReadMinMax)
cutSiteAnalysis("CPD", "all_reps", "1h", "NHF1", NHF1GeneralReadSequenceDataDirectory,
                c("CAC","CAG","CAT"), "CAB", humanReadMinMax)
cutSiteAnalysis("6-4", "all_reps", "20min", "NHF1", NHF1GeneralReadSequenceDataDirectory,
                c("CG","CA","TG","TA"), "YR", humanReadMinMax)
cutSiteAnalysis("cisplatin", "rep1", "90min", "GM12878", GM12878GeneralReadSequenceDataDirectory,
                c("CG","CA","TG","TA"), "YR", humanReadMinMax)
cutSiteAnalysis("CPD", "all_reps", "30m", "Dm", DmGeneralReadSequenceDataDirectory,
                c("CG","CA","TG","TA"), "YR", drosophilaReadMinMax)
cutSiteAnalysis("CPD", "all_reps", "1h", "NHF1", NHF1GeneralReadSequenceDataDirectory,
                c("TCAGTC","TCAGTT"), "TCAGTY", humanReadMinMax)
cutSiteAnalysis("CPD", "all_reps", "30m", "Dm", DmGeneralReadSequenceDataDirectory,
                c("TCAGTC","TCAGTT"), "TCAGTY", drosophilaReadMinMax)
```

>>>>

deaminationDataDirectory = file.path(bioinformaticsDirectory, "deamination_determination", "data")
NHF1GeneralReadSequenceDataDirectory = file.path(deaminationDataDirectory, "NHF1", "general_read_sequences")
GM12878GeneralReadSequenceDataDirectory = file.path(deaminationDataDirectory, "GM12878", "general_read_sequences")
DmGeneralReadSequenceDataDirectory = file.path(deaminationDataDirectory, "Dm", "general_read_sequences")
arabidopsisGeneralReadSequenceDataDirectory = file.path(deaminationDataDirectory, "Arabidopsis", "general_read_sequences")
yeastGeneralReadSequenceDataDirectory = file.path(deaminationDataDirectory, "yeast", "general_read_sequences")
EColiGeneralReadSequenceDataDirectory = file.path(deaminationDataDirectory, "E_coli_WT", "general_read_sequences")

humanReadMinMax = c(22, 30)
drosophilaReadMinMax = c(23, 32)
arabidopsisReadMinMax = c(24, 30)
yeastReadMinMax = c(22, 25)
EColiReadMinMax = c(11, 13)
```

@@

```{r define_all_that_and_a_bag_of_chips_function}
cutSiteAnalysis = function(lesion, repetition, timepoint, cellType, directory,
                           queryPattern, sequencesName, readMinMax,
                           dataTypeString = dataTypeStrings$alignedReads, fileExtension = ".bed") {
  readData = fread(generateFilePath(
    directory, dataTypeString, fileExtension,
    cellType = cellType, lesion = lesion, timepoint = timepoint,
    repitition = repetition, expansionNum = expansionOffset, filtering = filtering
  ))
  readData = data.table(Read_Length = nchar(readData$V7)-expansionOffset*2, Read_Sequence = readData$V7)
  readData = readData[Read_Length >= readMinMax[1] & Read_Length <= readMinMax[2]]

  seqFreq = getSequenceFreqByReadLengthAndPos(readData, queryPattern = queryPattern, posType = FIVE_PRIME)
  seqFreq[, Position := Position - 10]
  fwrite(seqFreq, sep = '\t', file = generateFilePath(
      directory = directory, dataTypeStrings$sequenceFrequency, ".tsv",
      cellType = cellType, lesion = lesion, repitition = repetition,
      timepoint = timepoint, sequence = sequencesName, filtering = filtering))
}
```

```{r compute_and_write_sequence_frequencies}
HYV = c("ACA","ACC","ACG","ATA","ATC","ATG",
        "CCA","CCC","CCG","CTA","CTC","CTG",
        "TCA","TCC","TCG","TTA","TTC","TTG")
HYVPattern = "^[ACT][CT][ACG]"

cutSiteAnalysis("CPD", "all_reps", "1h", "NHF1", NHF1GeneralReadSequenceDataDirectory,
                HYVPattern, "HYV", humanReadMinMax)
cutSiteAnalysis("6-4", "all_reps", "20min", "NHF1", NHF1GeneralReadSequenceDataDirectory,
                HYVPattern, "HYV", humanReadMinMax)
cutSiteAnalysis("cisplatin", "rep1", "90min", "GM12878", GM12878GeneralReadSequenceDataDirectory,
                HYVPattern, "HYV", humanReadMinMax)
cutSiteAnalysis("CPD", "all_reps", "20min", "yeast", yeastGeneralReadSequenceDataDirectory,
                HYVPattern, "HYV", yeastReadMinMax)
cutSiteAnalysis("CPD", "all_reps", "ZT2", "Arabidopsis", arabidopsisGeneralReadSequenceDataDirectory,
                HYVPattern, "HYV", arabidopsisReadMinMax)
cutSiteAnalysis("CPD", "all_reps", "5min", "E_coli_WT", EColiGeneralReadSequenceDataDirectory,
                HYVPattern, "HYV", EColiReadMinMax)
```

### 2022-07-30 R/5'CutSiteAnalysisFigureGeneration.rmd
HYV is better than YR!
Lots of changes to accommodate the change in sequence pattern from YR to HYV:
- Changed sequence frequency generation to except regular expression patterns.
- Updated various plotting and data generation notebooks to generate new HYV plots. This also included updating figure generation.

deaminationDataDirectory = file.path(bioinformaticsDirectory, "deamination_determination", "data")
NHF1GeneralReadSequenceDataDirectory = file.path(deaminationDataDirectory, "NHF1", "general_read_sequences")
GM12878GeneralReadSequenceDataDirectory = file.path(deaminationDataDirectory, "GM12878", "general_read_sequences")
DmGeneralReadSequenceDataDirectory = file.path(deaminationDataDirectory, "Dm", "general_read_sequences")

humanReadMinMax = c(22, 30)
drosophilaReadMinMax = c(23, 32)

@@

```{r define_all_that_and_a_bag_of_chips_function}
cutSitePlotting = function(lesion, repetition, timepoint, cellType, directory,
                           sequencesName, readMinMax, plottingPositionsFromEnd) {
  sequenceFrequencyData = fread(generateFilePath(
    directory, dataTypeStrings$sequenceFrequency, ".tsv",
    cellType = cellType, lesion = lesion, timepoint = timepoint,
    repitition = repetition, sequence = sequencesName, filtering = filtering
  ))

  sequenceFrequencyData[, Position := Position + 10]
  endPositionsByTimepointAndRL = sequenceFrequencyData[, .(End = max(abs(Position))),
                                                        by = Read_Length]
  setkey(endPositionsByTimepointAndRL, Read_Length)
  endPosByFullFreqRow = endPositionsByTimepointAndRL[list(sequenceFrequencyData$Read_Length)]$End
  seqFreqPlottingRows = sequenceFrequencyData[abs(Position) + abs(plottingPositionsFromEnd) > endPosByFullFreqRow]

  plotSequenceFrequencies(seqFreqPlottingRows,
                          title = paste(cellType, lesion, titleCaseFiltering, timepoint),
                          secondaryYAxisLabel = "Read Length", yAxisLabel = paste(sequencesName,"Frequency"),
                          showFivePrimeCutSite = TRUE, querySequences = c("CG","CA","TG","TA"), expansionOffset = 10,
                          minReadLength = readMinMax[1], maxReadLength = readMinMax[2], endBasedBackgroundNum = 8,
                          endBasedCheckPositions = -10, displayPValue = TRUE, xAxisLabel = "3' Relative Position")
}
```

```{r sequence_frequency_figures, fig.width = 5, fig.height = 7}
cutSitePlotting("CPD", "all_reps", "1h", "NHF1", NHF1GeneralReadSequenceDataDirectory,
                "YR", humanReadMinMax, 17)
cutSitePlotting("6-4", "all_reps", "20min", "NHF1", NHF1GeneralReadSequenceDataDirectory,
                "YR", humanReadMinMax, 17)
cutSitePlotting("CPD", "all_reps", "1h", "NHF1", NHF1GeneralReadSequenceDataDirectory,
                "CAB", humanReadMinMax, 17)
cutSitePlotting("cisplatin", "rep1", "90min", "GM12878", GM12878GeneralReadSequenceDataDirectory,
                "YR", humanReadMinMax, 17)
cutSitePlotting("CPD", "all_reps", "30m", "Dm", DmGeneralReadSequenceDataDirectory,
                "YR", drosophilaReadMinMax, 17)
cutSitePlotting("CPD", "all_reps", "1h", "NHF1", NHF1GeneralReadSequenceDataDirectory,
                "TCAGTY", humanReadMinMax, 25)
cutSitePlotting("CPD", "all_reps", "30m", "Dm", DmGeneralReadSequenceDataDirectory,
                "TCAGTY", drosophilaReadMinMax, 25)
```

>>>>

deaminationDataDirectory = file.path(bioinformaticsDirectory, "deamination_determination", "data")
NHF1GeneralReadSequenceDataDirectory = file.path(deaminationDataDirectory, "NHF1", "general_read_sequences")
NHF1ExcisionSiteDataDirectory = file.path(deaminationDataDirectory, "NHF1", "excision_site_analysis")
GM12878GeneralReadSequenceDataDirectory = file.path(deaminationDataDirectory, "GM12878", "general_read_sequences")
DmGeneralReadSequenceDataDirectory = file.path(deaminationDataDirectory, "Dm", "general_read_sequences")
arabidopsisGeneralReadSequenceDataDirectory = file.path(deaminationDataDirectory, "Arabidopsis", "general_read_sequences")
yeastGeneralReadSequenceDataDirectory = file.path(deaminationDataDirectory, "yeast", "general_read_sequences")
EColiGeneralReadSequenceDataDirectory = file.path(deaminationDataDirectory, "E_coli_WT", "general_read_sequences")

humanReadMinMax = c(22, 30)
drosophilaReadMinMax = c(23, 32)
arabidopsisReadMinMax = c(24, 30)
yeastReadMinMax = c(22, 25)
EColiReadMinMax = c(11, 13)

@@

```{r define_all_that_and_a_bag_of_chips_function}
cutSitePlotting = function(lesion, repetition, timepoint, cellType, directory,
                           sequencesName, readMinMax, maxPlottingPos, xAxisOffset = 1,
                           backgroundPositions = -9:-4, checkPositions = -1,
                           displayPValue = TRUE) {
  sequenceFrequencyData = fread(generateFilePath(
    directory, dataTypeStrings$sequenceFrequency, ".tsv",
    cellType = cellType, lesion = lesion, timepoint = timepoint,
    repitition = repetition, sequence = sequencesName, filtering = filtering
  ))

  plotSequenceFrequencies(sequenceFrequencyData[Position <= maxPlottingPos],
                          title = paste(cellType, lesion, titleCaseFiltering, timepoint),
                          secondaryYAxisLabel = "Read Length", yAxisLabel = paste(sequencesName,"Frequency"),
                          showFivePrimeCutSite = TRUE, querySequences = c("333"), expansionOffset = 10,
                          minReadLength = readMinMax[1], maxReadLength = readMinMax[2],
                          backgroundPositions = backgroundPositions, checkPositions = checkPositions,
                          displayPValue = displayPValue, xAxisLabel = "5' Relative Position",
                          xAxisOffset = xAxisOffset, xAxisBreaks = c(-8, 0, 8))
}
```

```{r sequence_frequency_figures, fig.width = 5, fig.height = 7}
cutSitePlotting("CPD", "all_reps", "1h", "NHF1", NHF1GeneralReadSequenceDataDirectory,
                "HYV", humanReadMinMax, 7)
cutSitePlotting("6-4", "all_reps", "20min", "NHF1", NHF1GeneralReadSequenceDataDirectory,
                "HYV", humanReadMinMax, 7)
cutSitePlotting("cisplatin", "rep1", "90min", "GM12878", GM12878GeneralReadSequenceDataDirectory,
                "HYV", humanReadMinMax, 7)
cutSitePlotting("CPD", "all_reps", "20min", "yeast", yeastGeneralReadSequenceDataDirectory,
                "HYV", yeastReadMinMax, 7)
cutSitePlotting("CPD", "all_reps", "ZT2", "Arabidopsis", arabidopsisGeneralReadSequenceDataDirectory,
                "HYV", arabidopsisReadMinMax, 7)
cutSitePlotting("CPD", "all_reps", "5min", "E_coli_WT", EColiGeneralReadSequenceDataDirectory,
                "HYV", EColiReadMinMax, 7, 1, NULL, NULL, FALSE)
```

### 2022-07-30 R/NHF1DeaminationAnchoredExcisionSiteDataGeneration.rmd
HYV is better than YR!
Lots of changes to accommodate the change in sequence pattern from YR to HYV:
- Changed sequence frequency generation to except regular expression patterns.
- Updated various plotting and data generation notebooks to generate new HYV plots. This also included updating figure generation.

>>>>

threePrimeMMAnchoredCPDSeqFreq = lapply(threePrimeMismatchAnchoredCPDData, function(mismatchDataRep) {
  lapply(mismatchDataRep, function(mismatchData) {
    getSequenceFreqByReadLengthAndPos(mismatchData, queryPattern = "^[ACT][CT][ACG]")
  })
})
for (i in seq_along(threePrimeMMAnchoredCPDSeqFreq)) {
  for (j in seq_along(threePrimeMMAnchoredCPDSeqFreq[[i]])) {
    fwrite(threePrimeMMAnchoredCPDSeqFreq[[i]][[j]], sep = '\t', file = generateFilePath(
      directory = deaminationExcisionSiteDataDirectory, dataTypeStrings$sequenceFrequency, ".tsv",
      cellType = cellType, lesion = "CPD", repitition = names(threePrimeMMAnchoredCPDSeqFreq)[i],
      timepoint = names(threePrimeMMAnchoredCPDSeqFreq[[i]])[j], strandPolarity = "three_prime",
      anchored = TRUE, sequence = "HYV", filtering = filtering
    ))
  }
}

### 2022-07-30 R/NHF1TandemDeaminationDataGeneration.rmd
HYV is better than YR!
Lots of changes to accommodate the change in sequence pattern from YR to HYV:
- Changed sequence frequency generation to except regular expression patterns.
- Updated various plotting and data generation notebooks to generate new HYV plots. This also included updating figure generation.

>>>>

threePrimeMMAnchoredCPDSeqFreq = lapply(threePrimeMismatchAnchoredCPDData, function(mismatchDataRep) {
  lapply(mismatchDataRep, function(mismatchData) {
    getSequenceFreqByReadLengthAndPos(mismatchData, queryPattern = "^[ACT][CT][ACG]")
  })
})
for (i in seq_along(threePrimeMMAnchoredCPDSeqFreq)) {
  for (j in seq_along(threePrimeMMAnchoredCPDSeqFreq[[i]])) {
    fwrite(threePrimeMMAnchoredCPDSeqFreq[[i]][[j]], sep = '\t', file = generateFilePath(
      directory = deaminationExcisionSiteDataDirectory, dataTypeStrings$sequenceFrequency, ".tsv",
      cellType = cellType, lesion = "CPD", repitition = names(threePrimeMMAnchoredCPDSeqFreq)[i],
      timepoint = names(threePrimeMMAnchoredCPDSeqFreq[[i]])[j], strandPolarity = "three_prime",
      anchored = TRUE, sequence = "HYV", filtering = filtering, includedMismatches = mismatchType
    ))
  }
}
